---
title: "For Developers"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggcircos)
library(ggplot2)
```

# Philosophy

There are two core components of this package.

1. An internal S7 class (CircosBuilder) built for parsing, storing, manipulating, querying and transforming the multiomic data. 
2. An exposed, highly opinionated api that lazily building up `ggcircos` plots.

For now, we plot the circos itself in ggplot2, and make use of many of its geoms and polar coordinate system, but thats about it. We handle the organisation of different layers of the circos, padding between reference karyotypes and coordinate transformation to respect inter-karyotype padding ourselves. This is important if we ever want to add additional graphing front-ends like base R. 

## The circos coordinate system

You can think of building a circos plot by first building up a series of vertically stacked rectangular layes that ggplot will later map to a polar coordinate system.

```{r}
gg_horizontal_rectangles <- ggplot() +
  annotate(geom = "rect", xmin = 1, xmax = 100, ymin = 1, ymax = 40, color = "black", fill = "grey") +
  annotate(geom = "rect", xmin = 1, xmax = 100, ymin = 60, ymax = 100, color = "black", fill = "pink") +
  annotate(geom = "text", x = 50, y = 80, label = "zone 1\n(y=60-100, x = 1-100)", color = "black") +
  annotate(geom = "text", x = 50, y = 20, label = "zone 2\n(y=1-40, x = 1-100)", color = "black") +
  theme_minimal() +
  theme(axis.text.y = element_blank())

# Normal coordinate system
gg_horizontal_rectangles

# Polar coordinates
gg_horizontal_rectangles + coord_radial(expand = FALSE)

# If you want axis labels to rotate perpendicular to circle tangent
gg_horizontal_rectangles + coord_radial(expand = FALSE) + guides(theta = guide_axis_theta(angle = 90), r = guide_axis())
```
But how do we turn this into a real circos-like structure? Well lets decide some rules for ourselves.

1. The canvas size is 100x100 and is 1-based start & inclusive. This means we should never allow any graphics we draw to leave 1-100 on either axis. 
2. We'll call each layer a 'zone'
3. Once polar coord system is applied, the, zones with higher ymin/ymax will be closer to the outer perimeter.
4. The outermost zone is reserved for a 'reference' backbone. 
5. Based on the total length (+ total gaps between refs) two functions are generated. `project_points` which maps chr:pos positions to their corresponding positions between 1-100 on the x axis, and `project_intervals` which does the same for chr:start-end intervals
6.  `project_intervals` will be applied to reference reference segments.


